Early Observations of the Type Ia Supernova \abc:\\ \textbf{a Case of Interaction with Nearby, Unbound Material and/or} Strong Ejecta Mixing

### 2017-11-23 doc/ms.tex
Update abstract slightly

Early observations of Type Ia supernovae (SNe) provide a unique probe of their
progenitor systems and explosion physics. Here, we report the intermediate
Palomar Transient Factory (iPTF) discovery of an extraordinarily young SN Ia,
\abc. By fitting a power law to our early light curve, we infer that first
light for the SN, \textbf{that is when the SN produces appreciable optical
radiation following shock breakout}, \textbf{occurred only}
$0.15\pm_{0.07}^{0.15}$\,d before our first detection. In the $\sim$24\,hr
after discovery, \abc\ rose by $\sim$2\,mag, \textbf{featuring} a near-linear
rise in flux for $\gtrsim$3 d. \textbf{Early spectra show strong \ion{C}{2}
absorption, which disappears after $\sim$7\,d.} Unlike the
extensively-observed Type Ia SN\,2011fe, the $(B-V)_0$ colors of \abc\ are
blue and nearly constant in the days after explosion. We show that our early
observations of \abc\ cannot be explained by either SN shock breakout and the
associated, subsequent cooling, or the SN ejecta colliding with a stellar
companion. Instead, we argue that the early characteristics of \abc,
including: (i) the rapid, near-linear rise, (ii) the non-evolving blue colors,
and (iii) the strong \textbf{\ion{C}{2}} absorption, are the result of either
\textbf{ejecta interaction with nearby, unbound material or vigorous mixing of
radioactive $^{56}$Ni in the SN ejecta,} or a combination of the two. In the
next few years, dozens of very young \textit{normal} SNe Ia will be
discovered, and observations similar to those presented here will constrain
the white dwarf explosion mechanism.

>>>>

Early observations of Type Ia supernovae (SNe) provide a unique probe of their
progenitor systems and explosion physics. Here, we report the intermediate
Palomar Transient Factory (iPTF) discovery of an extraordinarily young SN Ia,
\abc. By fitting a power law to our early light curve, we infer that first
light for the SN, \textbf{that is when the SN could have first been detected
by our survey}, \textbf{occurred only} $0.15\pm_{0.07}^{0.15}$\,d before our
first detection. In the $\sim$24\,hr after discovery, \abc\ rose by
$\sim$2\,mag, \textbf{featuring} a near-linear rise in flux for $\gtrsim$3 d.
\textbf{Early spectra show strong \ion{C}{2} absorption, which disappears
after $\sim$7\,d.} Unlike the extensively-observed Type Ia SN\,2011fe, the
$(B-V)_0$ colors of \abc\ are blue and nearly constant in the days after
explosion. We show that our early observations of \abc\ cannot be explained by
either SN shock breakout and the associated, subsequent cooling, or the SN
ejecta colliding with a stellar companion. Instead, we argue that the early
characteristics of \abc, including: (i) the rapid, near-linear rise, (ii) the
non-evolving blue colors, and (iii) the strong \textbf{\ion{C}{2}} absorption,
are the result of either \textbf{ejecta interaction with nearby, unbound
material or vigorous mixing of radioactive $^{56}$Ni in the SN ejecta,} or a
combination of the two. In the next few years, dozens of very young
\textit{normal} SNe Ia will be discovered, and observations similar to those
presented here will constrain the white dwarf explosion mechanism.
